---
title: "ElasticNet baseline"
output:
  BiocStyle::html_document:
      number_sections: yes
      toc: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Example for Classification Data -- Breast Invasive Carcinoma}
  %\VignetteEncoding{UTF-8}
params: 
  seed: !r 1985 * 2
  alpha: !r .7
  train.percentage: !r .7
  n.cores: !r 10
  #
  normalize: 'center' # max, center, none
  log2: !r TRUE
  foldid: !r TRUE
  network: 'string'
  network.cutoff: !r 0.1
  network.trans: 'hub' # hub, orphan, none
---

# Libraries

```{r, message=FALSE, warning=FALSE}
library(glmnet)
library(reshape2)
library(survival)
library(tidyverse)
library(Vennerable)
library(glmSparseNet)
library(loose.rock)
library(futile.logger)

```

```{r, include=FALSE}
devtools::load_all()

.Last.value <- base.dir(path = '/ssd_home/averissimo/work/rpackages/network.cox-cache')
.Last.value <- show.message(FALSE)
.Last.value <- flog.layout(layout.format('[~l] ~m'))
.Last.value <- flog.appender(appender.tee('logger.txt'))
theme_set(theme_minimal())
```


# Parameters

```{r, echo=FALSE}
show.params(params)
```

# Data

```{r load.data}
my.data <- run.cache(prepare.tcga.survival.data.old, 
                     'brca', 
                     'primary.solid.tumor', 
                     #
                     #input.type                = 'rna',
                     normalization             = params$normalize,
                     log2.pre.normalize        = params$log2,
                     # include.negative.survival = FALSE,
                     handle.duplicates         = 'keep_first', 
                     coding.genes              = TRUE,
                     #
                     cache.prefix              = 'tcga-data',
                     show.message = TRUE)
#
clinical <- my.data$clinical
#
xdata     <- my.data$xdata
ydata     <- my.data$ydata
xdata.raw <- my.data$xdata.raw
#
#
xdata.digest.cache     <- my.data$xdata.digest
xdata.raw.digest.cache <- my.data$xdata.raw.digest
ydata.digest.cache     <- my.data$ydata.digest
#
rm(my.data)
```

## Sample distribution of 5 variables

*note*: This is a fixed result, regardless of seed parameter in report.

```{r, echo=FALSE}
set.seed(987654321)
sample(seq(ncol(xdata)), 9) %>% { xdata[,.] } %>% melt %>%
  ggplot() + 
    geom_density(aes(value, color = X2)) + 
    #geom_freqpoly(aes(value, color = Var2), bins = 100) + 
    facet_wrap(~ X2, ncol = 3, scales = 'free') +
    theme(legend.position = 'none')

set.seed(987654321)
sample(seq(ncol(xdata)), 9) %>% { xdata[,.] } %>% melt %>%
  ggplot() + 
    #geom_density(aes(value, color = Var2)) + 
    geom_freqpoly(aes(value, color = X2), bins = 100) + 
    facet_wrap(~ X2, ncol = 3, scales = 'free') +
    theme(legend.position = 'none')
```

## Setup training and test sets

```{r}
set.seed(params$seed)
ixs        <- balanced.train.and.test(which(ydata$status), which(!ydata$status), train.perc = params$train.percentage)
xdata.test <- xdata[ixs$test,]
ydata.test <- ydata[ixs$test,]
#
xdata.train <- xdata[ixs$train,]
ydata.train <- ydata[ixs$train,]

xdata.train.digest <- glmSparseNet::digest.cache(xdata.train)
```

## Network to use in penalization

```{r, echo=FALSE, warning=FALSE, message=FALSE}
## Using Network or Not
#
#
#

# default using string from degree.type
network <- params$network
if (params$network == 'oldie') {
  #
  # Load old matrix used in SPARSA conference
  my.env <- new.env()
  load(sprintf('/ssd_home/averissimo/work/rpackages/brca.analysis/data/degree-%.6f.RData', .4), envir = my.env)
  network <- my.env$degree
  
} else if (params$network == 'string') {
  #
  # Load degree of STRING network downloaded from external sources
  network <- run.cache(string.db.homo.sapiens) %>% {
    run.cache(build.string.network, ., use.names = 'ensembl')
    } %>% {
      .@x[.@x <= params$network.cutoff] <- 0
      if (TRUE) {
        .@x <- (.@x != 0) * 1
      }
      Matrix::colSums(.) + Matrix::rowSums(.)
    }
  
  #
  # Adds missing genes as nodes with 0 degree
  
  network <- colnames(xdata.raw)[!colnames(xdata.raw) %in% names(network)] %>% {
   c(network, array(0, length(.), dimnames = list(.))) 
  } %>% { .[colnames(xdata.raw)] }
} else if (params$network == 'string.pedro') {
  #
  # Load degree of STRING network downloaded from external sources
  my.env <- new.env()
  load('/ssd_home/averissimo/work/rpackages/network.cox-big.files/saves/degree-string.RData', envir = my.env)
  ix.names <- colnames(xdata) %in% names(my.env$degree_uw)
  network <- my.env$degree_uw[colnames(xdata)]
}

# display Network
if (is.character(network)) {
  flog.info('Network to be used: %s', network)
} else {
  qplot(network, geom = 'freqpoly', bins = 100) + ggtitle(sprintf('Degree of nodes in %s network', params$network))
}
```

```{r, include=FALSE}
# Transformation function for degree before used as a penalty

if (params$network.trans == 'none') {
  network.trans <- function(x) { return(x) }
} else if (params$network.trans == 'hub') {
  network.trans <- glmSparseNet::hub.heuristic
} else if (params$network.trans == 'hub') {
  network.trans <- glmSparseNet::orphan.heuristic
} else {
  stop('Error: network.trans parameter not recognized.')
}
```

# Estimating models

## Model with `standardize = FALSE`

```{r fit1}
set.seed(params$seed)
if (params$foldid) {
  foldid <- balanced.cv.folds(ydata.train$status)$output
} else {
  foldid <- sample(rep(seq(10), length = nrow(ydata.train)))
}
set.seed(params$seed)
fit.train.nosd <- run.cache(cv.glmSparseNet.mclapply, xdata.train, Surv(ydata.train$time, ydata.train$status), 
                            network = network,
                            family  = 'cox', alpha = params$alpha, 
                            #
                            network.options = network.options.default(trans.fun  = network.trans,
                                                                      min.degree = .2,
                                                                      n.cores    = params$n.cores,
                                                                      cutoff     = params$network.cutoff),
                            #
                            foldid       = foldid, 
                            mc.cores     = params$n.cores,
                            standardize  = FALSE,
                            cache.digest = list(xdata.train.digest))
```

## Model with `standardize = TRUE`

```{r fit2}
set.seed(params$seed)
if (params$foldid) {
  foldid <- balanced.cv.folds(ydata.train$status)$output
} else {
  foldid <- sample(rep(seq(10), length = nrow(ydata.train)))
}
set.seed(params$seed)
fit.train.sd <- run.cache(cv.glmSparseNet.mclapply, xdata.train, Surv(ydata.train$time, ydata.train$status), 
                          network = network,
                          family  = 'cox', alpha = params$alpha, 
                          #
                          network.options = network.options.default(trans.fun  = network.trans,
                                                                    min.degree = .2,
                                                                    n.cores    = params$n.cores, 
                                                                    cutoff     = params$network.cutoff),
                          #
                          foldid       = foldid, 
                          mc.cores     = params$n.cores,
                          standardize  = TRUE,
                          cache.digest = list(xdata.train.digest))
```

# Results

## Description of models (`standardize = FALSE`)

```{r, echo=FALSE}
plot(fit.train.nosd)
```

## Description of models (`standardize = FALSE`)

```{r, echo=FALSE}
plot(fit.train.sd)
```

## Best models

```{r}
best.model.nosd <- coef(fit.train.nosd, s = 'lambda.min')[,1] %>% { .[.!=0] }
best.model.sd   <- coef(fit.train.sd, s = 'lambda.min')[,1] %>% { .[.!=0] }
```

## Selected features *(count)*

```{r, echo=FALSE}
c(No.Standadize = length(best.model.nosd), Standardize = length(best.model.sd))
```

Overlapping features/genes

```{r, echo=FALSE}
if(length(best.model.nosd) * length(best.model.sd) * 1 > 0) {
  vv <- Venn(list(standardize    = names(best.model.nosd),
                  non.standarize = names(best.model.sd))) %>% compute.Venn(type = 'squares')
  gp <- VennThemes(vv)
  plot(vv, gp = gp) 
} else {
  flog.warn('One of the models did not select any features, see section "Selected features"')
}
```

## Kaplan-meier curves

### Train

```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot.list <- list()
if(length(best.model.nosd) > 0) {
  plot.list <- c(plot.list, list(separate2GroupsCox(best.model.nosd, xdata.train[, names(best.model.nosd)], ydata.train, plot.title = 'standardize = FALSE')$plot))
} else {
  flog.warn('Model without standardize does not select any variables')
}
  
if(length(best.model.sd) > 0) {
  plot.list <- c(plot.list, list(separate2GroupsCox(best.model.sd, xdata.train[, names(best.model.sd)], ydata.train, plot.title = 'standardize = TRUE')$plot))
} else {
  flog.warn('Model with standardize does not select any variables')
}
   
if (length(plot.list) > 0) multiplot(plotlist = plot.list, ncol = 2)
```

### Test

```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot.list <- list()
if(length(best.model.nosd) > 0) {
  plot.list <- c(plot.list, list(separate2GroupsCox(best.model.nosd, xdata.test[, names(best.model.nosd)], ydata.test, plot.title = 'standardize = FALSE')$plot))
} else {
  flog.warn('Model without standardize does not select any variables')
}
  
if(length(best.model.sd) > 0) {
  plot.list <- c(plot.list, list(separate2GroupsCox(best.model.sd, xdata.test[, names(best.model.sd)], ydata.test, plot.title = 'standardize = TRUE')$plot))
} else {
  flog.warn('Model with standardize does not select any variables')
}
   
if (length(plot.list) > 0) multiplot(plotlist = plot.list, ncol = 2)
```



















```{r, eval=FALSE, include=FALSE}
#
# Use to generate multiple reporst
#
render.me <- function(my.list, default.list = list(alpha            = .7, 
                                                   foldid           = FALSE,
                                                   n.cores          = 10, 
                                                   seed             = 1985, 
                                                   train.percentage = .7, 
                                                   log2             = FALSE,
                                                   normalize        = 'none',
                                                   network          = 'none', 
                                                   network.cutoff   = 0,
                                                   network.trans    = 'none')) {
  
  my.list <- taRifx::merge.list(my.list, default.list)
  
  lapply(c('center', 'max', 'none'), function(ix) {
    return(tibble(normalize = c(ix, ix), log2 = c(TRUE,FALSE)))
  }) %>% 
  melt(id.vars = c('normalize', 'log2')) %>% 
  dplyr::select(1:2) %>% {
    for(line.ix in seq(nrow(.))) {
      taRifx::merge.list(list(normalize = .[line.ix, 'normalize'], log2 = .[line.ix, 'log2']), my.list) %>% {
        
        .$foldid <- TRUE
        output_file <- sprintf('baseline_reports/network_%s--cutoff_%.2f--trans_%s--normalize_%s--log2_%s--foldid_%s--seed_%d.html', 
                               .$network, .$network.cutoff, .$network.trans, .$normalize, .$log2, .$foldid, .$seed)
        
        tryCatch(rmarkdown::render('baseline.Rmd', params = ., output_file = output_file))
        
        # .$foldid <- FALSE
        # output_file <- sprintf('baseline_reports/network_%s--cutoff_%.2f--trans_%s--normalize_%s--log2_%s--foldid_%s--seed_%d.html', 
        #                        .$network, .$network.cutoff, .$network.trans, .$normalize, .$log2, .$foldid, .$seed)
        # tryCatch(rmarkdown::render('baseline.Rmd', params = ., output_file = output_file))
      }
    }
  }
} 

for(ix.seed in c(1985*2, 1985)) {
  # tryCatch(render.me(list(seed = ix.seed, network = 'none')))
  tryCatch(render.me(list(seed = ix.seed, network.trans = 'hub',  network = 'string')))
  # tryCatch(render.me(list(seed = ix.seed, network.trans = 'orphan', network = 'string')))
  
  #tryCatch(render.me(list(seed = ix.seed, network.trans = 'hub',  network = 'correlation', network.cutoff = .6)))
  #tryCatch(render.me(list(seed = ix.seed, network.trans = 'hub',  network = 'covariance', network.cutoff = .1)))
  #tryCatch(render.me(list(seed = ix.seed, network.trans = 'hub',  network = 'string.pedro')))
  
  #tryCatch(render.me(list(seed = ix.seed, network.trans = 'orphan', network = 'correlation', network.cutoff = .6)))
  #tryCatch(render.me(list(seed = ix.seed, network.trans = 'orphan', network = 'covariance', network.cutoff = .1)))
  #tryCatch(render.me(list(seed = ix.seed, network.trans = 'orphan', network = 'string.pedro')))
}
```
