---
title: "Cross-Validation discovery"
author: "André Veríssimo"
date: "`r Sys.Date()`"
output:
  github_document:
    fig_width: 10
  html_document:
    toc: true
    self_contained: true
    number_sections: true
    fig_width: 10
  BiocStyle::html_document:
    number_sections: yes
    toc: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Example for Classification Data -- Breast Invasive Carcinoma}
  %\VignetteEncoding{UTF-8}
params:
  project: 'brca' # skcm prad brca
  tissue: 'primary.solid.tumor' # primary.solid.tumor metastatic solid.tissue.normal
  coding.genes: !r TRUE
  degree.unweighted: !r TRUE
  degree.correlation: 'pearson'
  degree.cutoff: !r 0
  #
  degree.type: "string" # covariance, correlation, correlation.inv, oldie, sparsebn, string
  #
  handle.duplicates: 'keep_first' # keep_all , keep_first
  #
  # glmnet parameter
  alpha: !r seq(0, 1, .1)
  # subset of variables to be used (Inf for all)
  subset: !r Inf
  #
  seed: !r 1985
  n.cores: !r 10
  #
  calc.params.old: !r FALSE
---

## Libraries

```{r, messages=FALSE, warning=FALSE}
library(futile.logger)
library(parallel)
library(glmnet)
library(loose.rock)
library(digest)
library(ggplot2)
library(reshape2)
library(survival)
library(brca.data)
library(Vennerable)
library(limma)
library(tidyverse)
library(forcats)
library(survival.owl)
library(biclust)
#
library(glmSparseNet)
```

```{r, include=FALSE}
devtools::install_github('averissimo/glmnet')
#
devtools::load_all()
#
.Last.value <- base.dir(path = '/ssd_home/averissimo/work/rpackages/network.cox-cache')
.Last.value <- show.message(FALSE)
.Last.value <- flog.layout(layout.format('[~l] ~m'))
.Last.value <- flog.appender(appender.tee('logger.txt'))
theme_set(theme_minimal())
```

## Parameters

```{r parms, echo=FALSE}
max.chars <- max(sapply(names(params), nchar))
for (ix.names in sort(names(params))) {
  prefix <- paste(array(' ', max.chars - nchar(ix.names)), collapse = '')
  if (is.vector(params[[ix.names]]) && length(params[[ix.names]]) == 1) {
    if (is.list(params[[ix.names]])) {
      flog.info('  %s%s: %s', prefix, ix.names, paste(params[[ix.names]], collapse = ', '))
    }
    else if (is.character(params[[ix.names]])) {
      flog.info('  %s%s: %s', prefix, ix.names, params[[ix.names]])
    }  else if (is.infinite(params[[ix.names]])) {
      flog.info('  %s%s: %f', prefix, ix.names, params[[ix.names]])
    } else if (is.integer(params[[ix.names]])) {
      flog.info('  %s%s: % 11d', prefix, ix.names, params[[ix.names]])
    } else {
      flog.info('  %s%s: % 11.3f', prefix, ix.names, params[[ix.names]])
    }
  } else if (is.vector(params[[ix.names]])) {
    flog.info('  %s%s: %s', prefix, ix.names, paste(params[[ix.names]], collapse = ', '))
  } else {
    flog.info('  %s%s: (i do not know how to display this)', prefix, ix.names)
  }
}
```

# Load and normalize data

## Load TCGA data

```{r, eval=FALSE}
# xdata     - matrix with FPKM gene level expression
# ydata     - survival data
# xdata.raw - original matrix with possible multiple tissue samples from same indiviudal
prepare.tcga.survival.data(params$project, 
                           params$tissue, 
                           #
                           handle.duplicates            = params$handle.duplicates, 
                           center                       = TRUE,
                           log2.normalize               = TRUE,
                           include.negative.survival    = TRUE,
                           input.type                   = 'rna',
                           #
                           cache.prefix = 'tcga-data',
                           coding.genes = params$coding.genes)
```


```{r load.data, include=FALSE}
my.data <- run.cache(prepare.tcga.survival.data, 
                     params$project, 
                     params$tissue, 
                     #
                     handle.duplicates            = params$handle.duplicates, 
                     center                       = TRUE,
                     log2.normalize               = TRUE,
                     include.negative.survival    = TRUE,
                     input.type                   = 'rna',
                     #
                     cache.prefix = 'tcga-data',
                     coding.genes = params$coding.genes)
#
clinical <- my.data$clinical
#
xdata     <- my.data$xdata
ydata     <- my.data$ydata
xdata.raw <- my.data$xdata.raw
#
#
xdata.digest.cache     <- my.data$xdata.digest
xdata.raw.digest.cache <- my.data$xdata.raw.digest
ydata.digest.cache     <- my.data$ydata.digest
#
rm(my.data)
```

## Load degree data

```{r degree.load}
if (params$degree.type == 'oldie') {
  #
  # Load old matrix used in SPARSA conference
  load(sprintf('/home/averissimo/work/rpackages/brca.analysis/data/degree-%.6f.RData', params$degree.cutoff))
  
} else if (params$degree.type == 'correlation') {
  #
  # Load degree of network calculated summing the inverse of each weight
  degree <- degree.cor(xdata.raw, 
                       consider.unweighted = params$degree.unweighted, 
                       cutoff              = params$degree.cutoff, 
                       method              = params$degree.correlation,
                       n.cores = params$n.cores)
} else if (params$degree.type == 'covariance') {
  #
  # Load degree of network calculated using covariance
  degree <- degree.cov(xdata.raw, 
                       consider.unweighted = params$degree.unweighted, 
                       cutoff              = params$degree.cutoff, 
                       method              = params$degree.correlation,
                       n.cores = params$n.cores)
} else if (params$degree.type == 'sparsebn') {
  #
  # Load degree of network from sparsebn package (calulate bayesian network)
  degree <- degree.sparsebn(xdata.raw,
                            cutoff              = params$degree.cutoff, 
                            consider.unweighted = params$degree.unweighted,
                            n.cores = params$n.cores)
} else if (params$degree.type == 'string') {
  #
  # Load degree of STRING network downloaded from external sources
  degree <- run.cache(string.db.homo.sapiens) %>% {
    run.cache(build.string.network, ., use.names = 'ensembl')
    } %>% {
      .@x[.@x <= params$degree.cutoff] <- 0
      if (params$degree.unweighted) {
        .@x <- (.@x != 0) * 1
      }
      Matrix::colSums(.) + Matrix::rowSums(.)
    }
  
  #
  # Adds missing genes as nodes with 0 degree
  
  degree <- colnames(xdata.raw)[!colnames(xdata.raw) %in% names(degree)] %>% {
   c(degree, array(0, length(.), dimnames = list(.))) 
  } %>% { .[colnames(xdata.raw)] }
}
```

### Penalty

## Preparing degree vector

* Normalize degree between 0 and 1
* DegreeCox: 1 - degree
* OrphanCox: degree
* `trans.fun` is a double power to scale the values

see `?glmSparseNet::heuristic.scale` or `?glmSparseNet::hub.heuristic`

```{r transformation_function}
# see ?glmSparseNet::heuristic.scale
trans.fun <- function(x) { heuristic.scale(x, min.ret = .2) }
```


```{r prepare.penalty, echo=FALSE}
original.penalty.factor        <- degree
names(original.penalty.factor) <- colnames(xdata)

# if (params$add.age.at.diag) {
#   #
#   age.at.diag.norm <- c(xdata.train[,'age.at.diag'],
#                         xdata.test[,'age.at.diag'])
#   age.at.diag.norm <- age.at.diag.norm[colnames(xdata)]
#   age.cov <- unlist(mclapply(xdata.ix.no.added, function(ix) {
#     co.exp <- cov(age.at.diag.norm, xdata[ix,], method = 'pearson')
#     if (abs(co.exp) >= params$degree.perct)
#       return(co.exp)
#     else {
#       return(0)
#     }
#   }, mc.cores = params$n.cores))
#   original.penalty.factor <- original.penalty.factor + age.cov
#   original.penalty.factor <- c(original.penalty.factor, age.at.diag = sum(age.cov))
# }

##########################
#                        #
#   SUPER IMPORTANT!!!!  #
#                        #
##########################
original.penalty.factor[is.na(original.penalty.factor)] <- 0
norm.orig.penalty.factor <- original.penalty.factor / max(original.penalty.factor[!is.na(original.penalty.factor)])

#
#
# DegreeCox (old and log(old))
#
penalty.factor.degree.log <- penalty.factor.degree.old <- 1 / norm.orig.penalty.factor

inf.ix <- is.infinite(penalty.factor.degree.old) # index for infinite values

# log(old)
log.penal                          <- log(penalty.factor.degree.old[!inf.ix]) + 1
penalty.factor.degree.log[!inf.ix] <- log.penal
penalty.factor.degree.log[inf.ix]  <- max(log.penal) + 1

# old
non.log                           <- penalty.factor.degree.old[!inf.ix]
penalty.factor.degree.old[inf.ix] <- max(non.log) + 1

#
# DegreeCox heuristic
#
penalty.factor.degree.new <-trans.fun(1 - norm.orig.penalty.factor)

#
# OrphanCox
#
penalty.factor.orphan <- trans.fun(norm.orig.penalty.factor)
```

## Cross-validation

Testing classic elastic-net, DegreeCox and OrphanCox with cross-validation to find optimal number of variables for comparisson models.

```{r cv.models, echo=FALSE}
alpha.results <- alpha.km <- list()

for (ix.alpha in params$alpha) {
  cv.classic <- my.cv.glmnet('cv.classic', 
                             ix.alpha,
                             xdata,
                             ydata,
                             rep(1, ncol(xdata)), 
                             lambda.min.ratio = .001,
                             n.cores          = params$n.cores,
                             seed             = params$seed)
  
  cv.degree  <- my.cv.glmnet('cv.degree', 
                             ix.alpha,
                             xdata,
                             ydata,
                             penalty.factor.degree.new, 
                             lambda.min.ratio = .001,
                             n.cores          = params$n.cores,
                             seed             = params$seed)
  cv.orphan  <- my.cv.glmnet('cv.degree', 
                             ix.alpha,
                             xdata,
                             ydata,
                             penalty.factor.orphan, 
                             lambda.min.ratio = .001,
                             n.cores          = params$n.cores,
                             seed             = params$seed)  
  
  alpha.results[[sprintf('%.3f', ix.alpha)]] <- list(glmnet = sum(coef(cv.classic$model, s = 'lambda.min')[,1] != 0),
                                                     degree = sum(coef(cv.degree$model, s = 'lambda.min')[,1] != 0),
                                                     orphan = sum(coef(cv.orphan$model, s = 'lambda.min')[,1] != 0))
  
  alpha.km[[sprintf('%.3f', ix.alpha)]] <- list(glmnet = my.draw.kaplan(coef(cv.classic$model, s = 'lambda.min')[,1], '', xdata, ydata)$pvalue,
                                                degree = my.draw.kaplan(coef(cv.degree$model, s = 'lambda.min')[,1], '', xdata, ydata)$pvalue,
                                                orphan = my.draw.kaplan(coef(cv.orphan$model, s = 'lambda.min')[,1], '', xdata, ydata)$pvalue)
}
```

```{r}
melt(alpha.km) %>% 
  mutate(Alpha = L1, Model = L2) %>%
  filter(Alpha != '0.000') %>% 
  ggplot() + geom_point(aes(x = Alpha, y = value, color = Model), alpha = .5)
```

```{r}
melt(alpha.results) %>% 
  mutate(Alpha = L1, Model = L2) %>%
  filter(Alpha != '0.000') %>% 
  ggplot() + geom_point(aes(x = Alpha, y = value, color = Model))
```



```{r, eval=FALSE}
show.cv.info <- function(my.model, name) {
  selected.genes <- sort(colnames(xdata)[my.model$coef != 0])
  flog.info(' * %s: %d -- %s', name, sum(my.model$coef != 0), paste(selected.genes, collapse = ', '))
  return(selected.genes)
}

flog.info('10-fold cross-validation')
cv.classic.genes <- show.cv.info(cv.classic, 'classic')
cv.degree.genes  <- show.cv.info(cv.degree, ' degree')
cv.orphan.genes  <- show.cv.info(cv.orphan, ' orphan')

tryCatch({
vv <- Venn(list(Classic = cv.classic.genes, Degree = cv.degree.genes, Orphan = cv.orphan.genes))
plot(vv, doWeights = FALSE)
}, error = function(err) { flog.error('Probel with %s', err)})
#
km.cv <- list()

km.cv[['cv.classic']] <- draw.kaplan(list(CV.Classic = cv.classic$coef), xdata[,xdata.ix], ydata, plot.title = 'CV.Classic', legend.outside = F)
km.cv[['cv.degree']] <- draw.kaplan(list(CV.Degree = cv.degree$coef), xdata[,xdata.ix], ydata, plot.title = 'CV.Degree', legend.outside = F)
km.cv[['cv.orphan']] <- draw.kaplan(list(CV.Orphan = cv.orphan$coef), xdata[,xdata.ix], ydata, plot.title = 'CV.Orphan', legend.outside = F)

for (ix in names(km.cv)) {
  print(km.cv[[ix]]$plot)
}
```

