---
title: "Cross-Validation discovery"
author: "André Veríssimo"
date: "`r Sys.Date()`"
output:
  github_document:
    fig_width: 10
  html_document:
    toc: true
    self_contained: true
    number_sections: true
    fig_width: 10
    dev: svg
  BiocStyle::html_document:
    number_sections: yes
    toc: true
    dev: svg
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Example for Classification Data -- Breast Invasive Carcinoma}
  %\VignetteEncoding{UTF-8}
params:
  project: 'brca' # skcm prad brca
  tissue: 'primary.solid.tumor' # primary.solid.tumor metastatic solid.tissue.normal
  coding.genes: !r TRUE
  degree.unweighted: !r TRUE
  degree.correlation: 'pearson'
  degree.cutoff: !r 0
  #
  degree.type: "string" # covariance, correlation, correlation.inv, oldie, sparsebn, string
  #
  normalize: 'center' # max, center, none
  log2: !r TRUE
  #
  handle.duplicates: 'keep_first' # keep_all , keep_first
  #
  # glmnet parameter
  alpha: !r seq(.1, 1, .1)
  # subset of variables to be used (Inf for all)
  subset: !r Inf
  #
  seed: !r 1985
  seed.additional: !r 1
  n.cores: !r 10
  #
  calc.params.old: !r FALSE
---

```{r, include=FALSE, eval=FALSE}
rmarkdown::render('Analysis.Rmd', output_file = 'brca-center.html', params = list(normalization = 'center'), output_format = BiocStyle::html_document())
rmarkdown::render('OptimalVariableSize.Rmd', output_format = BiocStyle::html_document())
rmarkdown::render('Analysis.Rmd', output_file = 'brca-max.html', params = list(normalization = 'max'), output_format = BiocStyle::html_document())
```


# Libraries

```{r, message=FALSE, warning=FALSE}
library(futile.logger)
library(parallel)
library(glmnet)
library(loose.rock)
library(digest)
library(ggplot2)
library(reshape2)
library(survival)
library(brca.data)
library(Vennerable)
library(limma)
library(tidyverse)
library(forcats)
library(survival.owl)
library(biclust)
#
library(glmSparseNet)

require(doMC)
registerDoMC(cores=params$n.cores)

#devtools::install_github('averissimo/glmnet')
#
devtools::load_all()
#
.Last.value <- base.dir(path = '/ssd_home/averissimo/work/rpackages/network.cox-cache')
.Last.value <- show.message(FALSE)
.Last.value <- flog.layout(layout.format('[~l ~t] ~m'))
.Last.value <- flog.appender(appender.tee(sprintf('logger-%s.txt', format(Sys.time(), '%Y%m%d-%Hh%Mm%S'))))
theme_set(theme_minimal())
```

# Load and normalize data

## Load TCGA data

```{r load.data, include=FALSE}
my.data <- run.cache(prepare.tcga.survival.data.old, 
                     params$project, 
                     params$tissue, 
                     #
                     #input.type                = 'rna',
                     normalization             = params$normalize,
                     log2.pre.normalize        = params$log2,
                     #include.negative.survival = FALSE,
                     #
                     handle.duplicates         = params$handle.duplicates, 
                     coding.genes              = params$coding.genes,
                     cache.prefix              = 'tcga-data',
                     show.message = T)
#
clinical <- my.data$clinical
#
xdata     <- my.data$xdata
ydata     <- my.data$ydata
xdata.raw <- my.data$xdata.raw
#
#
xdata.digest.cache     <- my.data$xdata.digest
xdata.raw.digest.cache <- my.data$xdata.raw.digest
ydata.digest.cache     <- my.data$ydata.digest
#
rm(my.data)
```

## Load degree data

```{r degree.load}
if (params$degree.type == 'oldie') {
  #
  # Load old matrix used in SPARSA conference
  load(sprintf('/home/averissimo/work/rpackages/brca.analysis/data/degree-%.6f.RData', params$degree.cutoff))
  
} else if (params$degree.type == 'correlation') {
  #
  # Load degree of network calculated summing the inverse of each weight
  degree <- degree.cor(xdata.raw, 
                       consider.unweighted = params$degree.unweighted, 
                       cutoff              = params$degree.cutoff, 
                       method              = params$degree.correlation,
                       n.cores = params$n.cores)
} else if (params$degree.type == 'covariance') {
  #
  # Load degree of network calculated using covariance
  degree <- degree.cov(xdata.raw, 
                       consider.unweighted = params$degree.unweighted, 
                       cutoff              = params$degree.cutoff, 
                       method              = params$degree.correlation,
                       n.cores = params$n.cores)
} else if (params$degree.type == 'sparsebn') {
  #
  # Load degree of network from sparsebn package (calulate bayesian network)
  degree <- degree.sparsebn(xdata.raw,
                            cutoff              = params$degree.cutoff, 
                            consider.unweighted = params$degree.unweighted,
                            n.cores = params$n.cores)
} else if (params$degree.type == 'string') {
  #
  # Load degree of STRING network downloaded from external sources
  degree <- run.cache(string.db.homo.sapiens) %>% {
    run.cache(build.string.network, ., use.names = 'ensembl')
    } %>% {
      .@x[.@x <= params$degree.cutoff] <- 0
      if (params$degree.unweighted) {
        .@x <- (.@x != 0) * 1
      }
      Matrix::colSums(.) + Matrix::rowSums(.)
    }
  
  #
  # Adds missing genes as nodes with 0 degree
  
  degree <- colnames(xdata.raw)[!colnames(xdata.raw) %in% names(degree)] %>% {
   c(degree, array(0, length(.), dimnames = list(.))) 
  } %>% { .[colnames(xdata.raw)] }
} else if (params$degree.type == 'string.pedro') {
  #
  # Load degree of STRING network downloaded from external sources
  load('/ssd_home/averissimo/work/rpackages/network.cox-big.files/saves/degree-string.RData')
  ix.names <- colnames(xdata) %in% names(degree_uw)
  degree <- degree_uw[colnames(xdata)]
}
```

### Penalty

## Preparing degree vector

* Normalize degree between 0 and 1
* DegreeCox: 1 - degree
* OrphanCox: degree
* `trans.fun` is a double power to scale the values

see `?glmSparseNet::heuristic.scale` or `?glmSparseNet::hub.heuristic`

```{r transformation_function}
# see ?glmSparseNet::heuristic.scale
trans.fun <- function(x) { heuristic.scale(x) + 0.2}
```


```{r prepare.penalty, echo=FALSE}
original.penalty.factor        <- degree
names(original.penalty.factor) <- colnames(xdata)

##########################
#                        #
#   SUPER IMPORTANT!!!!  #
#                        #
##########################
original.penalty.factor[is.na(original.penalty.factor)] <- 0
norm.orig.penalty.factor <- original.penalty.factor / max(original.penalty.factor[!is.na(original.penalty.factor)])

#
#
# DegreeCox (old and log(old))
#
penalty.factor.degree.log <- penalty.factor.degree.old <- 1 / norm.orig.penalty.factor

inf.ix <- is.infinite(penalty.factor.degree.old) # index for infinite values

# log(old)
log.penal                          <- log(penalty.factor.degree.old[!inf.ix]) + 1
penalty.factor.degree.log[!inf.ix] <- log.penal
penalty.factor.degree.log[inf.ix]  <- max(log.penal) + 1

# old
non.log                           <- penalty.factor.degree.old[!inf.ix]
penalty.factor.degree.old[inf.ix] <- max(non.log) + 1

#
# DegreeCox heuristic
#
penalty.factor.degree.new <-trans.fun(1 - norm.orig.penalty.factor)

#
# OrphanCox
#
penalty.factor.orphan <- trans.fun(norm.orig.penalty.factor)
```

## Cross-validation

Testing classic elastic-net, DegreeCox and OrphanCox with cross-validation to find optimal number of variables for comparisson models.

```{r cv.models, echo=FALSE}
alpha.models <- tibble(alpha = numeric(), seed = numeric(), glmnet = list(), hub = list(), orphan = list())

for (ix.seed in seq(params$seed, params$seed + params$seed.additional)) {
  flog.info('Starting run with seed %d', ix.seed)
  for (ix.alpha in params$alpha) {
    #aux.call <- function(xdata, ydata, network, alpha, seed) { my.cv.glmnet(alpha, xdata, ydata, network, parallel = TRUE, seed = seed) }
    aux.call <- function(xdata, ydata, network, alpha, seed) { my.cv.glmnet(alpha, xdata, ydata, network, parallel = params$n.cores, seed = seed) }
    
    cv.classic <- run.cache(aux.call, xdata, ydata, rep(1, ncol(xdata)), ix.alpha, ix.seed, 
                            cache.prefix = 'cv.glmSparseNet.elastic.net.glmnet', cache.digest = list(xdata.digest.cache))
    flog.info('    - Finished elastic net (%d)', sum(cv.classic$coef != 0))

    cv.hub     <- run.cache(aux.call, xdata, ydata, penalty.factor.degree.new, ix.alpha, ix.seed, 
                            cache.prefix = 'cv.glmSparseNet.hub.glmnet', cache.digest = list(xdata.digest.cache))
    flog.info('    - Finished hub (%d)', sum(cv.hub$coef != 0))
    
    cv.orphan  <- run.cache(aux.call, xdata, ydata, penalty.factor.orphan, ix.alpha, ix.seed, 
                            cache.prefix = 'cv.glmSparseNet.orphan.glmnet', cache.digest = list(xdata.digest.cache))
    flog.info('    - Finished orphan (%d)', sum(cv.orphan$coef != 0))
    
    new.line <- tibble(alpha  = ix.alpha, 
                       seed   = ix.seed,
                       glmnet = list(cv.classic$model), 
                       hub    = list(cv.hub$model), 
                       orphan = list(cv.orphan$model))
    
    alpha.models <- rbind(alpha.models, new.line)
    
    save(new.line, file = sprintf('cv.value/cv.value_%s_xdata-%s_ydata-%s', glmSparseNet::digest.cache(new.line), xdata.digest.cache, loose.rock::digest.cache(ydata)))
    
    flog.info('  * Finished run with alpha %.3f', ix.alpha)
  }
}
```

## Selected variables

```{r, echo=FALSE}
sel.all <- apply(alpha.models[,-2], 1:2, function(ix.name) {
  if (is.numeric(ix.name[[1]])) {
    return(ix.name[[1]])
  } else {
    best.coef <- coef(ix.name[[1]], s = 'lambda.min')[,1]
    return(names(best.coef[best.coef != 0]))
  }
}) %>% as.data.frame %>% as.tbl

sel.all$alpha <- as.numeric(sel.all$alpha)

sel.all %>% as.data.frame %>% group_by(alpha) %>% summarise_all(funs(mean = mean(sapply(., length)),
                                                                      sd = sd(sapply(., length)),
                                                                      unique = length(unique(unlist(.))))) %>% knitr::kable()
```

## Kaplan-Meier Results per `alpha` parameter

Individuals are separated by the coefficients estimated by the model in two groups with high and low risk of the event *(death)*. The median is then used.

### Summary table for each model type *(mean is used)*

```{r, echo=FALSE}
km.all <- apply(alpha.models[,-2], 1:2, function(ix.name) {
  if (is.numeric(ix.name[[1]])) {
    return(ix.name[[1]])
  } else {
    best.coef <- coef(ix.name[[1]], s = 'lambda.min')[,1]
    if (sum(best.coef[best.coef != 0]) == 0) {
      return(NA)
    }
    return(draw.kaplan(best.coef, xdata, ydata)$pvalue)
  }
})

km.all %>% as.data.frame %>%
  group_by(alpha) %>% 
  summarise_all(funs(mean = mean(., na.rm = TRUE), 
                     SE = sd(., na.rm = TRUE))) %>% 
  knitr::kable()
```

### Kaplan-Meier estimator plot

```{r, echo=FALSE}
km.all %>% as.data.frame %>% 
  melt.data.frame(id.vars = 'alpha') %>% 
  mutate(Alpha = factor(alpha), Model = variable) %>%
  ggplot(aes(x = Alpha, y = value, color = Model)) + 
  geom_boxplot(alpha = .5) + 
  # geom_jitter(width = 0.2, size = .5) +
  #expand_limits(y = .001) +
  ggtitle('Kaplan-Meier estimator (per alpha parameter)', 
          subtitle = 'Two groups (High/Low risk of event) are estimated based on Cox model\'s coefficients')
```


## Concordance C-Index Results per `alpha` parameter

Concordance C-Index calculates a value between 0 and 1 that represents the pairwise concordance between the survival time and the relative-risk calculated from the Cox model. In other words, it assesses for each pair of individuals if individuals with lower risk survive longer than individuals with higher risks.

### Summary of results

```{r, echo=FALSE}
c.index.all <- apply(alpha.models[,-2], 1:2, function(ix.name) {
  if (is.numeric(ix.name[[1]])) {
    return(ix.name[[1]])
  } else {
    best.coef <- coef(ix.name[[1]], s = 'lambda.min')[,1]
    if (sum(best.coef[best.coef != 0]) == 0) {
      return(NA)
    }
    return(fit.risk(best.coef, xdata, ydata, NULL))
  }
})

c.index.all %>% as.data.frame %>%  group_by(alpha) %>% 
  summarise_all(funs(mean = mean(., na.rm = TRUE), 
                     SE   = sd(., na.rm = TRUE))) %>% 
  knitr::kable()
```

### C-Index plot

```{r}
c.index.all  %>% as.data.frame %>% 
  melt.data.frame(id.vars = 'alpha') %>% 
  mutate(Alpha = factor(alpha), Model = variable) %>%
  ggplot(aes(x = Alpha, y = value, color = Model)) + 
  geom_boxplot(alpha = .5) + 
  # geom_jitter(width = 0.2) +
  ggtitle('Concordance C-Index (per alpha parameter)')
```



## Parameters

```{r parms, echo=FALSE}
show.params(params)
```
